Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-01-08 paper/paper.tex
revision of section 7.1

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction, without taking the contraction order finding and just in time compiling time into account.
    Legends are properties, algebra and devices that we used in the computation, one can find the corresponding computed property in \Tbl{tbl:generictypes}.
    (a) time and space complexity versus the number of vertices for the benchmarked graphs.
    (b) The computing time for calculating the MIS size and for counting of the number of independent sets (ISs), the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for configuration enumeration, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    }
    \label{fig:benchmark}
\end{figure}

Figure (a) shows the time and space complexity of tensor network contraction for variety of graph sizes, where the space complexity is the same as the treewidth of the problem graph.
The contraction order is obtained using the local search algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique is used for graphs with treewidth greater than 27 to fit the computation into a 16GB memory.
One can see all the computing times in figure (b), (c) and (d) have a strong correlation with the treewidth.
%The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
%computing independence polynomial (panel c) and find one independent set and enumerating independent sets (panel d).
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions to evaluate tensor contractions, hence 
are much faster comparing to non-BLAS element types in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough such that the actual computing time is comparable or larger than the launching overhead of CUDA kernels.
Most algebras can be computed on GPU, except those requiring dynamic sized structures, i.e. \texttt{PN} and \texttt{SN}.
In figure (c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
however, it maybe suffer from the round off errors. The finite field (GF$(p)$) approach is the only method that does not have round-off errors and can run on a GPU.
In figure (d), one can see the technique to bound the enumeration space improves the performance for more than one order in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a device with 32GB main memory.

>>>>

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction, without taking the contraction order finding and just-in-time compiling time into account.
    Legends are properties, algebra and devices that we used in the computation; one can find the corresponding computed property in \Tbl{tbl:generictypes}.
    (a) time and space complexity versus the number of vertices for the benchmarked graphs.
    (b) The computing time for calculating the MIS size and for counting of the number of all independent sets (ISs), the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for configuration enumeration, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    }
    \label{fig:benchmark}
\end{figure}

Figure (a) shows the time and space complexity of tensor network contraction for different graph sizes, where the space complexity is the same as the treewidth of the problem graph.
The contraction order is obtained using the local search algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique is used for graphs with treewidth greater than 27 to fit the computation into a 16GB memory.
One can see that all the computing times in figure (b), (c) and (d) have a strong correlation with the treewidth.
%The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
%computing independence polynomial (panel c) and find one independent set and enumerating independent sets (panel d).
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions to evaluate tensor contractions, hence 
are much faster comparing to non-BLAS element types in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough such that the actual computing time is comparable or larger than the launching overhead of CUDA kernels.
Most algebras can be computed on GPU, except those requiring dynamic sized structures, i.e.\ \texttt{PN} and \texttt{SN}.
In figure (c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
but it may suffer from round-off errors. The finite field (GF$(p)$) approach is the only method that does not have round-off errors and can be run on a GPU.
In figure (d), one can see the technique to bound the enumeration space improves the performance for more than one order of magnitude in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a device with 32GB main memory.
